---
title: "Random Walk Chain"
output: html_document
---

Back to Example 7.2, Section 7.1.1

Suppose we observe $y_1, ..., y_n$ from the mixture density
$$
f(y | \delta) = \delta N(7, 0.5^2) + (1 - \delta) N(10, 0.5^2)
$$
We assume a uniform prior for $\delta$.  Then the posterior density is proportional to
$$
g(\delta) \propto \prod_{j=1}^n f(y_j | \delta).
$$

We'd like to use a random walk M-H chain to simulate a sample from the posterior.

First generate some data from mixture distribution and plot the data.

```{r}
n <- 20
delta <- 0.7
set.seed(1234)
y <- ifelse(runif(n) < delta, 
            rnorm(n, 7, 0.5),
            rnorm(n, 10, 0.5))
library(MASS)
truehist(y, 20, main="Histogram of observed data")
```

We define and graph the log posterior for of the transformed parameter $\theta = \log\left(\frac{\delta}{1-\delta}\right)$

```{r}
logpost = function(theta, y) {
  delta <- exp(theta) / (1 + exp(theta))
  sum(log(delta * dnorm(y, 7, 0.5) +
            (1 - delta) * dnorm(y, 10, 0.5))) + theta -
            2 * log(1 + exp(theta))
}
theta <- seq(-1, 5, by=0.01)
plot(theta, exp(sapply(theta, logpost, y)), type="l",
     ylab="Density", main="Graph of Posterior of theta")
```

Illustrate several choices for scale parameter of random walk algorithm.  (I am using the function rwmetrop in the LearnBayes package to do the sampling.)

First, I use a scale parameter of 0.05 and run the random walk algorithm for 10,000 iterations.  I'll output the acceptance rate and construct a trace plot and a density plot of the simulated draws.

```{r}
library(LearnBayes)
scale.value <- 0.05
out1 <- rwmetrop(logpost, 
                 list(var=1, scale=scale.value),
                 start=0,
                 m=10000,
                 y)
out1$accept
plot(mcmc(out1$par), main=paste("Scale =", scale.value))
```

Repeat this for 10,000 iterations using a scale parameter of 2.

```{r}
scale.value <- 2.0
out2 <- rwmetrop(logpost, 
                 list(var=1, scale=scale.value),
                 start=0,
                 m=10000,
                 y)
out2$accept
plot(mcmc(out2$par), main=paste("Scale =", scale.value))
```

